---
title: "week5"
author: "danbom"
date: '2020 10 2 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Base R

```{r}
suppressMessages(library(tidyverse))
```

## R Objects and Variables

- Everything in R is stored as an object, which is associated with a variable name.
- An object is a technical terminology defined in Object Oriented Programming (OOP). (OOP is an important concept but not in this class)
- A variable name can be assigned to an object using the assignment operator.

```{r}
# store a number to a variable named `a`
a <- 0.2
```

```{r}
# print a
a
```

```{r}
# store a vector to a variable named `b`
b <- c(1,4,9)
```

```{r}
# print b
b
```

```{r}
z <- 5
i <- (z * 2 + 45)/2
i
```

# Function

```{r}
# call (execute or run) the mean() function
mean(c(1,2,3,4))    
```

```{r}
# mean() will not work with NA
mean(c(1,2,NA,4))     
```

```{r}
# When na.rm = TRUE, NA will be removed before computation
mean(c(1,2,NA,4), na.rm = TRUE)    
```

## User-Defined Functions

- function.name <- function(arg1, arg2, arg3){body}

```{r}
# Define se() function that calculate the standard error
se <- function(x) {
  v <- var(x)
  n <- length(x)
  return(sqrt(v/n))
}
```

```{r}
mySample <- rnorm(n=100, mean=20, sd=4)
mySample # sd는 표준편차
se(mySample)
```

```{r}
x1 <- rnorm(100, mean=0, sd=1)  # generate 100 random numbers from Normal(0,1)
x2 <- rnorm(100, mean=3, sd=2)  # generate 100 random numbers from Normal(3,2)
```

## Vectors

- Type : typeof() returns the type of an object.
- Attributes: attributes() returns additional arbitrary metadata

```{r}
typeof(c(1,2,3))
length(c(1,2,3))
attributes(c(1,2,3))
```

## Attributes

```{r}
a <- c(x=1,y=2,z=3)   # when creating
names(a)
```

```{r}
names(a) <- c("l", "m", "n")   # by modifying existing names
a
```

```{r}
attributes(a)    # names are attributes
```

```{r}
attributes(mtcars)
```

## Type Coercion (Conversion)

```{r}
v <- c(4, 7, 23.5, 76.2, 80, "rrt", 90)
v
```

```{r}
typeof(v)
```

```{r}
sum(c(TRUE, FALSE, TRUE))
```

- You can explicitly convert data type with as.character(), as.double(), as.integer(), and as.logical().

```{r}
a <- c(1,2,3)
a
```

```{r}
b <- as.character(a)
b
```

## NA represents missing

```{r}
u <- c(4, 6, NA, 2)
u
```

```{r}
k <- c(TRUE, FALSE, FALSE, NA, TRUE)
k
```

## Generate a vector

```{r}
# c() also combine vectors
a <- c(1,2,3)
b <- c(4,5,6)
c <- c(a, b)
c
```


```{r}
# k:n generates a vector whose elements are the sequence of numbers from k to n
1:10
```

```{r}
# seq() generates regular sequence
# seq(from, to)
seq(1, 10)
```

```{r}
# seq(from, to, by)
seq(1, 10, 2)
```

```{r}
# rep(x, times) replicates the values in x multiple times
# x can be a number or vector
# replicates 1 5 times
rep(1, 5)
```

```{r}
# replicates c(1, 2) 5 times
rep(c(1,2), 5)
```

```{r}
# each element of c(1,2) is repeated 5 times
rep(c(1,2), each = 5)
```

```{r}
# rnorm(n, mean = 0, sd = 1) generates a vector of n random samples 
# from a normal distribution with specific mean and sd. 
rnorm(100)
```

```{r}
qplot(rnorm(10000))
```

```{r}
# runif(n, min, max) generates a vector of n random samples 
# from a uniform distribution whose limits are min and max. 
runif(100, 0, 1)
```

```{r}
qplot(runif(10000, 0, 1))
```

## Indexing or subsetting a Vector

- Positive integers return elements at the specified positions.

```{r}
x <- c(2,3,4,5,6,7)
x[c(3,1)]
```

- Negative integers omit elements at the specified positions:

```{r}
x[-c(3,1)]
```

```{r}
x[c(TRUE, TRUE, FALSE, FALSE, TRUE, TRUE)]
```

```{r}
x > 3
```

```{r}
# This is called a logical indexing, which is a very powerful tool.
# > : greater than (Logical Operators)
x[x > 3]   
```

```{r}
x[x > 3 & x < 5]
```

```{r}
# %in% operator 
# v1 %in% v2 returns a logical vector indicating 
# whether the elements of v1 are included in v2. 
c(1,2,3) %in% c(2,3,4,5,6)
```

```{r}
a <- c(1,2,3,4,5)
a
```

```{r}
a[c(1,5)] <- 100
a
```

```{r}
a <- c(1,2,3,NA,5,6,NA)
a
```

```{r}
# is.na indicates which elements are missing
is.na(a)  # returns TRUE when missing
```

```{r}
# Type conversion: TRUE and FALSE will be converted into 1 and 0, respectively. 
# This expression answers the question: How many NSs are in a?
sum(is.na(a))
```

```{r}
# !x = not x (negation)
!is.na(a)  # returns TRUE when not missing
```

```{r}
# This expression answers the question: How many non-NSs are in a?
sum(!is.na(a))
```

```{r}
# logical indexing
a[is.na(a)] <- 999
a
```

```{r}
# create a vector with names
a <- c(x = 1, y = 2, z = 3)
a
```

```{r}
# named vector can be indexed using their names
a[c("x", "z")]
```

```{r}
# R uses a "recycling rule" by repeating the shorter vector
# In this example, R recycled c(TRUE, FALSE) to produce c(TRUE, FALSE, TRUE, FALSE)
i <- c(TRUE, FALSE)
a <- c(1,2,3,4)
a[i]
```

```{r}
# R uses a "recycling rule" by repeating the shorter vector
v1 <- c(4,5,6,7)
v2 <- c(10,10)
v1+v2
```

## Arrange a vector

```{r}
# sort(x, decreasing = FALSE) 
# By default, sort() sorts ascending order.
sort(c(5,6,4))
```

```{r}
# sorts into descending order
sort(c(5,6,4), decreasing = TRUE)
```

```{r}
# rev() provides a reversed version of its argument
rev(c(5,6,4))
```

```{r}
# rank() returns the sample ranks of the elements in a vector
rank(c(5,6,4))
```

```{r}
# order() returns a permutation which rearranges 
# its first argument into ascending or descending order. 
# What this means is order(c(5,6,4)) 
# 1) first sorts a vector in ascending order to produce c(4,5,6)
# 2) and returns the indices of the sorted element in the original vector. 
# e.g., we have 3 first b/c the index of 4 in the original vector is 3
# e.g., we have 1 first b/c the index of 5 in the original vector is 1
# e.g., we have 2 first b/c the index of 6 in the original vector is 2
order(c(5,6,4))
```

```{r}
# We use order() to sort a vector or dataframe
a <- c(5,6,4)
a[order(a)]
```

```{r}
# sort a dataframe
head(mtcars[order(mtcars$mpg), ])
```

## Vectorization of Functions

```{r}
v <- c(1,4,25)
sqrt(v)
```

```{r}
v <- c(1,2,3)
v^2
```

```{r}
# mean will be subtracted from every element of v1
v1 <- c(1,2,3,4)
v1 - mean(v1)
```

## More functions

```{r}
# table() creates a frequency table
a <- c(1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4, 4, 4)
table(a)
```

```{r}
# unique() returns a vector of unique elements
unique(a)
```

```{r}
a <- c(1,2,3,NA,5)
```


```{r}
# na.rm = TRUE removes NAs before computation
mean(a, na.rm = TRUE)
```

## Generating Sequences

```{r}
5:0
```

```{r}
seq(from=1, to=3, by=0.5)
```

```{r}
rep(1:2, 3)
```

```{r}
# gl() generates sequences involving factors
# gl(k,n), k = the number of levels, 
# n = the number of repetitions.
gl(7,3)
```

## Exercise

```{r}
a <- mtcars$mpg
```

```{r}
length(a)
```

```{r}
# 평균구하기
sum(a)/length(a)
mean(a)
```

- Calculate the variance of a using sd() function.

```{r}
sd(a)^2
var(a)
sum((a-mean(a))^2)/(length(a)-1)
```

- Standardize the vector a

```{r}
(a-mean(a))/sd(a)
scale(a)
```

```{r}
max(a)-min(a)
# another solution
diff(range(a))
```

- Normalize the vector a

```{r}
# your maximum value will be 1, and minimum value will be 0. 
(a-min(a))/(max(a)-min(a))
```

```{r}
qplot(a)
```

```{r}
# To set aesthetics, wrap in I()
qplot(a, color = I("yellow"), fill = I("gray"))
```

```{r}
# creates a logical vector in which TRUE indicates the element that is larger than 20
a>20
```

```{r}
# This is a logical indexing where the logical vector 
# within the subsetting operator (i.e., []) will create a vector with elements larger than 20. 
a[a>20]
```

```{r}
length(a[a>20])
# same result
# this happens because of "vectorization" and "type conversion"
sum(a>20)
```

## Factors

- Factors are used to represent categorical data (e.g., gender, states).

## creating

```{r}
# levels are the set of allowed values
f1 <- factor(c(2,1,1,3,2,1,1))
f1
```

```{r}
typeof(f1)
```

```{r}
attributes(f1)
```

```{r}
levels(f1)
```

- The factor’s level is a character vector.

```{r}
# test for objects of type "character"
is.character(levels(f1))
```

- can change levels

```{r}
levels(f1) <- c("one", "two", "three")
f1
```

```{r}
f2 <- factor(c("Dec", "Apr", "Jan", "Mar"))
f2
```

```{r}
sort(f2)
```

```{r}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)
f3 <- factor(c("Dec", "Apr", "Jan", "Mar"), levels = month_levels)
f3
```

```{r}
# In many cases, this is the result that we expect. 
sort(f3)
```

```{r}
table(f3)
```

```{r}
table(f2)
```

```{r}
# the default level is in alphabetical order 
f4 <- factor(c("high", "low", "medium", "medium", "high"))
f4
```

```{r}
sort(f4)
```

```{r}
f5 <- factor(c("high", "low", "medium", "medium", "high"), levels = c("low", "medium", "high"))
f5
```

```{r}
sort(f5)
```

```{r}
f6 <- factor(c("high", "low", "medium", "medium", "high"), levels = c("low", "medium", "high"), ordered = TRUE)
f6
```

```{r}
min(f6)
```

```{r}
f6[1] > f6[2]
```

```{r}
f7 <- ordered(c("high", "low", "medium", "medium", "high"), levels = c("low", "medium", "high"))
f7
```

## Exercise

```{r}
x <- c("Agree", "Disagree", "Neutral", "Agree" ,"Agree", "Strongly disagree", "Neutral")
```

```{r}
# you may want to this
factor(x, levels = c("Strongly disagree", "Disagree", "Neutral",  "Agree", "Strongly agree"))
```

```{r}
x = c("male", "male", "female", "male", "female")
```


```{r}
# What I've asked you is to change the default alphabetical order using levels options. 
factor(x, levels = c("male", "female"))
```

```{r}
# I just wanted to introduce 'cut()` function
set.seed(7)
x <- rnorm(100)
cut(x, breaks = quantile(x))
```

## list

## creating

```{r}
x <- list(1:6, "a", c(TRUE, TRUE, FALSE), c(1.2, 3.3, 4.6, 6.6))
x
```

```{r}
# str() display the internal structure of an R object
str(x)
```

```{r}
typeof(x)
```

```{r}
# In R, lm() is a function that fits a regression model to data.
# In the following R expression, 'mpg' is a dependent variable
# and `disp` and `cyl` are independent variable. 
fit <- lm(mpg ~ disp + cyl, data = mtcars)
fit
```

```{r}
typeof(fit)
```

```{r}
str(fit)
```

## Subsetting a List

```{r}
my.lst <- list(stud.id=34453,      # creat a list
               stud.name="John", 
               stud.marks=c(14.3,12,15,19))
```

```{r}
my.lst
```

```{r}
# [ ] extracts a sub-list 
my.lst[1]
```

```{r}
typeof(my.lst[1])
```

```{r}
# [[ ]] extracts the value of an individual element 
my.lst[[1]]
```

```{r}
typeof(my.lst[[1]])
```

```{r}
# my.lst[[3]] will index the third element of a list, which is a numeric vector
# my.lst[[3]][2] will index the second element of the numeric vector
my.lst[[3]][2]
```

```{r}
# In the case of lists with named elements
# $ extracts the value of an individual element 
my.lst$stud.id
```

```{r}
typeof(my.lst$stud.id)
```

## Exercise

- fit is a list that contains the outputs of the lm() function for linear regression. Explore the structure of the fit object using str().

```{r}
fit <- lm(mpg ~ disp + cyl, data = mtcars)
fit
```

```{r}
# fit$coefficient is a vector
fit$coefficients
```

```{r}
# So, we can subset the first element using the following expression
fit$coefficients[1]
```

```{r}
# We can also use the name of element for indexing. 
fit$coefficients["(Intercept)"]
```

## Data frames

```{r}
head(iris)
```

```{r}
typeof(iris)
```

```{r}
str(iris)
```

## The apply family of functions

- The apply() family of functions refers to apply(), lapply(), sapply(), vapply(), mapply(), rapply(), and tapply().

- Why do we need them?
-- They will make your code much shorter by replacing your own copy and paste

```{r}
# A motivating example: check the number of missing values in each column of the following data frame 'm'
m <- data.frame(matrix(c(1,2,3,4,NA,6,7,NA,NA,NA,NA,NA), ncol = 4))
m
```

```{r}
sum(is.na(m$X1))
```

```{r}
sum(is.na(m$X2))
```

```{r}
sum(is.na(m$X3))
```

```{r}
sum(is.na(m$X4))
```

- lapply(X, FUN)
-- X = a list object in R
-- FUN = a function in R

- lapply()
-- takes a function (FUN)
-- applies it to each element of a list (X)
-- and returns the results in the form of a list

```{r}
# This one line of code will still work even when the number of columns are 1000 or more. 
lapply(m, function(x) sum(is.na(x)))
```

```{r}
# lapply() returns a list, whereas sapply() returns a vector, matrix, or array. 
sapply(m, function(x) sum(is.na(x)))
```

## Exercise

```{r}
# install.packages("psych")
library(psych)
```

```{r}
head(bfi)
```

## Matrices and Arrays

```{r}
a <- 1:6
```

```{r}
# Get or set specific attributes of an object.
attr(a, "dim") <- c(3,2)
a
```

```{r}
# by default, a matrix is filled by column
a <- matrix(1:6, ncol=3, nrow=2)
a
```

```{r}
a <- matrix(1:6, ncol=3)
a
```

```{r}
# a matrix can be filled by row using `byrow = TRUE`
a <- matrix(1:6, ncol=3, nrow=2, byrow = TRUE)
a
```

```{r}
# 2행 3열이라는 뜻
attributes(a)
```

```{r}
dim(a)
```

```{r}
b <- array(1:12, c(2,3,2))
b
```

```{r}
dim(b)
```

- length() generalises to nrow() and ncol() for matrices, and dim() for arrays.
- names() generalises to rownames() and colnames() for matrices, and dimnames(), a list of character vectors, for arrays.

```{r}
results <- matrix(c(10, 30, 40, 50, 43, 56, 21, 30), 2, 4, byrow = TRUE)
colnames(results) <- c("1qrt", "2qrt", "3qrt", "4qrt")
rownames(results) <- c("store1", "store2")
results
```

## Subsetting a Matrix and Array

```{r}
a <- matrix(1:9, nrow = 3)
colnames(a) <- c("A", "B", "C")
a
```

```{r}
a[c(TRUE, FALSE, TRUE), c("B", "A")]
```

```{r}
a[1, c(2,3)]
```

```{r}
# If you omit any dimension, you obtain full columns or rows
a[2,]
```

```{r}
a[,3]
```

```{r}
a > 3
```

```{r}
a[a>3] <- NA
a
```

## Combine Matrices by Columns or Rows

```{r}
# combine by columns
cbind(a,a)   
```

```{r}
# combine by rows
rbind(a,a)   
```
## Names of the Columns and Rows of Matrices

```{r}
colnames(a) 
```

```{r}
rownames(a) <- c("D", "E", "F")
rownames(a)
```

## Data Frames

```{r}
df <- data.frame(x=1:3, y=c("a", "b", "c"))
df
```

```{r}
# display the internal structure of an R object
str(df) 
```

- data.frame() converts strings into factors by default.
-- This default setting can cause serious problems in some cases.
-- stringAsFactors = FALSE suppresses this default setting.
-- Using str() to check data types is always a good practice.

```{r}
df <- data.frame(x=1:3, y=c("a", "b", "c"), stringsAsFactors = FALSE)
```

```{r}
str(df) # display the internal structure of an R object
```

## Control Flow

- Control flow = the order in which individual statement are executed (or evaluated)

## if-else

```{r}
x <- -5
if (x>0) {
  print("Positive number")
} else {
  print("Negative number")
}
```

## for

```{r}
for (i in 1:3) {
  print(i^2)
}
```

```{r}
# count the number of even numbers
x <- c(2,5,3,9,8,11,6)
count <- 0
for (val in x) {
  if(val %% 2 == 0)  count = count+1
}
print(count)
```

